---
title             : "Relationship between social capital and election results"
shorttitle        : "Title"

author: 
  - name          : "Anisha Babu"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    email         : "ababu@uoregon.edu"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - Conceptualization
      - Data Analysis
      - Writing - Original Draft Preparation
      - Writing - Review & Editing
      
  - name          : "Hyeonjin Cha"
    affiliation   : "1"
    email         : "hcha@uoregon.edu"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - Conceptualization
      - Data Analysis
      - Writing - Original Draft Preparation
      - Writing - Review & Editing
      
  - name          : "Diana DeWald"
    affiliation   : "1"
    email         : "ddewald@uoregon.edu"    
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - Conceptualization
      - Data Analysis
      - Writing - Original Draft Preparation
      - Writing - Review & Editing
      
  - name          : "Murat Kezer"
    affiliation   : "1"
    email         : "mkezer@uoregon.edu"    
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - Conceptualization
      - Data Analysis
      - Writing - Original Draft Preparation
      - Writing - Review & Editing               

affiliation:
  - id            : "1"
    institution   : "University of Oregon"


authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

  Enter author note here.

abstract: |
  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.
  
  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.
  
  One sentence clearly stating the **general problem** being addressed by  this particular study.
  
  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).
  
  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.
  
  One or two sentences to put the results into a more **general context**.
  
  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.
  
  <!-- https://tinyurl.com/ybremelq -->
  
keywords          : "keywords"
wordcount         : "X"

bibliography      : ["r-references.bib"]

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
library("papaja")
r_refs("r-references.bib")
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed,
                      eval = TRUE,
                      echo = FALSE,
                      cache = FALSE)
```

```{r packages}
library(rio); library(here); library(tidyverse); library(knitr); library(janitor)
```

```{r load data}
election <- import(here("data", "countypres_2000-2016.csv")) %>% 
  clean_names() #load election results

s_capital_97_04 <- import(here("data", "social-capital-variables-for-1997-2005-2009.xlsx")) %>% 
  clean_names() #load social capital data for 1997 through 2009

s_capital_09 <- import(here("data", "social_capital_2014.xlsx")) %>% 
  clean_names()#load social capital data for 2014
```

# Introduction
Social science literature has extensively examined the relationship between social capital and politics (e.g. Morales & Guigni, 2016; Jottier & Heyndels, 2012; La Due Lake & Huckfeldt, 1998). However, relatively little is known on the impact of social capital election results.

# Methods
We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study. <!-- 21-word solution (Simmons, Nelson & Simonsohn, 2012; retrieved from http://ssrn.com/abstract=2160588) -->

## Participants

## Material

## Procedure

## Data analysis
We used `r cite_r("r-references.bib")` for all our analyses.


# Results

# Discussion


\newpage

# References

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id="refs" custom-style="Bibliography"></div>
\endgroup
