---
title             : "Relationship between social capital and election results"
shorttitle        : "Title"
author: 
      
  - name          : "Anisha Babu"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    email         : "ababu@uoregon.edu"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - Conceptualization
      - Data Analysis
      - Writing - Original Draft Preparation
      - Writing - Review & Editing
  - name          : "Hyeonjin Cha"
    affiliation   : "1"
    email         : "hcha@uoregon.edu"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - Conceptualization
      - Data Analysis
      - Writing - Original Draft Preparation
      - Writing - Review & Editing
  - name          : "Diana DeWald"
    affiliation   : "1"
    email         : "ddewald@uoregon.edu"    
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - Conceptualization
      - Data Analysis
      - Writing - Original Draft Preparation
      - Writing - Review & Editing
  - name          : "Murat Kezer"
    affiliation   : "1"
    email         : "mkezer@uoregon.edu"    
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - Conceptualization
      - Data Analysis
      - Writing - Original Draft Preparation
      - Writing - Review & Editing               
affiliation:
  - id            : "1"
    institution   : "University of Oregon"
authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.
  Enter author note here.
abstract: |
  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.
  
  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.
  
  One sentence clearly stating the **general problem** being addressed by  this particular study.
  
  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).
  
  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.
  
  One or two sentences to put the results into a more **general context**.
  
  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.
  
  
  <!-- https://tinyurl.com/ybremelq -->
keywords          : "keywords"
wordcount         : "X"
bibliography      : ["r-references.bib"]
floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no
documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
library(papaja)
library(tidyverse)
library(rio)
library(knitr)
library(janitor)
library(magrittr)
library(stringr)
library(here)
r_refs("r-references.bib")
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed,
                      eval = TRUE,
                      echo = FALSE,
                      cache = FALSE)
```

# Data Preparation

## Load data and clean names
We first load the datasets and clean the variable names.
```{r load data}
election <- import(here("data", "countypres_2000-2016.csv")) %>% 
  clean_names() #load election results

s_capital_97 <- import(here("data", "social-capital-variables-for-1997.xlsx")) %>% 
  clean_names() #load social capital data for 1997

s_capital_05 <- import(here("data", "social-capital-variables-for-2005.xlsx")) %>% 
  clean_names() #load social capital data for 2005

s_capital_09 <- import(here("data", "social-capital-variables-for-2009.xlsx")) %>% 
  clean_names() #load social capital data for 2009

s_capital_14 <- import(here("data", "social-capital-variables-spreadsheet-for-2014.xlsx")) %>% 
  clean_names()#load social capital data for 2014

```

## Clean data
### Election data
- We start with the election data as it is more comprehensive in terms of the number of counties. First, we select the variables of interests. Then, we select the election year (i.e., 2000, 2008, 2012, 2016) that we will match with social capital data.  

- The name of the year variable is changed in a way that shows it is the year of election (so that it is not mixed with the same year variable in social capital data).  

- We create new datasets for each presidential election we are interested in. These will be later merged with corresponding social capital data.  

### Social capital data
- For each social capital dataset (i.e., 1997, 2005, 2009, 2014), we first add state code for some counties that do not readily contain that information. Then, we create two variables out of the area name such that we have different variables for county names and state codes.  

- We select the relevant variables and clean the variable names.  

- We create a year variable indicating when the data were collected.  

- Finally, we reorder variables so that the order of the variables is the same across datasets. This will be useful when we want to merge social capital data across year so that we can get descriptive statistics for each year simultaneously and that we can visualize the changes across years in social capital.  

```{r clean datasets}
#select, filter, and clean data
##election
election %<>%
  select(-office, -version, -candidate, -state) %>% #select relevant variables
  filter(year %in% c(2000, 2008, 2012, 2016)) #select relevant years

colnames(election)[1] <- "year_elctn" #rename year variable for merging

elctn_00 <- election %>% filter(year_elctn == 2000) #create a df for each election of interest so that it's easier to match them with sc dfs
elctn_08 <- election %>% filter(year_elctn == 2008)
elctn_12 <- election %>% filter(year_elctn == 2012)
elctn_16 <- election %>% filter(year_elctn == 2016)

##social capital 1997
s_capital_97[290, 2] <- "District of Columbia, DC" #add DC's code

s_capital_97 %<>% #create county and state variables
  separate(areaname, c("county","state_po"), sep = ', ')

s_capital_97 %<>% #select relevant variables
  select(-fitnes97, -memspt97, -memnec97)

colnames(s_capital_97)[4:18] <- str_sub(colnames(s_capital_97)[4:18], end=-3) #remove the year from the variable names

s_capital_97 %<>% #create year variable before merging
  mutate(year_sc = 1997)

##social capital 2005
s_capital_05[290, 2] <- "District of Columbia, DC" #add DC's code

s_capital_05 %<>% #create county and state variables
  separate(areaname, c("county","state_po"), sep = ', ')

s_capital_05 %<>% #select relevant variables
  select(-fitns05)

colnames(s_capital_05)[4:18] <- str_sub(colnames(s_capital_05)[4:18], end=-3) #remove the year from the variable names

s_capital_05 <- s_capital_05[,c(1:3, 10, 5, 11, 4, 12, 7:8, 6, 9, 15, 14, 13, 16, 17:18)] #reorder the variables

s_capital_05$year_sc <- 2005 #create year variable before merging

##social capital 2009
s_capital_09[290, 2] <- "District of Columbia, DC" #add DC's code

s_capital_09 %<>% #create county and state variables
  separate(areaname, c("county","state_po"), sep = ', ')

s_capital_09 %<>% #select relevant variables
  select(-fitns09)

colnames(s_capital_09)[4:18] <- str_sub(colnames(s_capital_09)[4:18], end=-3) #remove the year from the variable names

s_capital_09 <- s_capital_09[,c(1:3, 10, 5, 11, 4, 12, 7:8, 6, 9, 14:15, 13, 16:18)] #reorder the variables

s_capital_09 %<>% #create year variable before merging
  mutate(year_sc = 2009)

##social capital 2014
s_capital_14[77, 2] <- "Hoonah-Angoon Census Area, AK" #remove the repeated state code

s_capital_14 %<>% #create county and state variables
  separate(county_name, c("county","state_po"), sep = ', ') 

s_capital_14$county <- word(s_capital_14$county, 1) #remove the word 'county' and keep only the name

s_capital_14 %<>% #select relevant variables
  select(-recreational2014)

colnames(s_capital_14)[4:18] <- str_sub(colnames(s_capital_14)[4:18], end=-5) #remove the year from the variable names

s_capital_14 <- s_capital_14[, c(1:3, 10, 5, 11, 4, 12, 7:8, 6, 9, 16, 15, 13, 17, 14, 18)] #reorder the variables

s_capital_14 %<>% #create year variable before merging
  mutate(year_sc = 2014)

colnames(s_capital_14) <- colnames(s_capital_09) #rename variables to make it compatible with previous years' datasets
```

## Merge Datasets

- First, we merge social capital data across years for reasons explained above, and call it `s_capital`.  

- Next, we merge corresponding election and social capital data for 4 time points. **In doing so,** we keep the rows that exist in both election and social capital data. For instance, if we do not have the election information for a county, we do not include it in the merged dataset even if we have that county's social capital data. These datasets are called `df_year`. *Year* denotes the year of election. Also, we remove the duplicate variables (i.e., state and county names) and fix the names. We did not remove them earlier because we first wanted to merge the social capital data with all the variables.  

- Finally, we merge all election and social capital data in the same dataset (i.e., `df`). 

```{r merge datasets}
s_capital <- bind_rows(s_capital_97, s_capital_05, s_capital_09, s_capital_14) #merge social capital data

df_00 <- inner_join(elctn_00, s_capital_97, by = "fips") #merge election 2000 & social capital 1997
df_00 <- df_00[, c(1, 25, 2:7, 10:24)] #remove duplicates and reorder variables
colnames(df_00)[3:4] <- c("state_po", "county") #rename variables

df_08 <- inner_join(elctn_08, s_capital_05, by = "fips") #merge election 2008 & social capital 2005
df_08 <- df_08[, c(1, 25, 2:7, 10:24)] #remove duplicates and reorder variables
colnames(df_08)[3:4] <- c("state_po", "county") #rename variables

df_12 <- inner_join(elctn_12, s_capital_09, by = "fips") #merge election 2012 & social capital 2009
df_12 <- df_12[, c(1, 25, 2:7, 10:24)] #remove duplicates and reorder variables
colnames(df_12)[3:4] <- c("state_po", "county") #rename variables

df_16 <- inner_join(elctn_16, s_capital_14, by = "fips") #merge election 2016 & social capital 2014
df_16 <- df_16[, c(1, 25, 2:7, 10:24)] #remove duplicates and reorder variables
colnames(df_16)[3:4] <- c("state_po", "county") #rename variables

df <- bind_rows(df_00, df_08, df_12, df_16) #merged dataset: contains all election and corresponding social capital 


```



# Introduction
Social science literature has extensively examined the relationship between social capital and politics (e.g. Morales & Guigni, 2016; Jottier & Heyndels, 2012; La Due Lake & Huckfeldt, 1998). However, relatively little is known on the impact of social capital election results.

# Methods
We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study. <!-- 21-word solution (Simmons, Nelson & Simonsohn, 2012; retrieved from http://ssrn.com/abstract=2160588) -->

## Participants

## Material

## Procedure

## Data analysis
We used `r cite_r("r-references.bib")` for all our analyses.

```{r descriptives set up, echo = FALSE, results = "hide"}
#install.packages("kableExtra")
library(kableExtra)

(table_01 <- df %>% 
  group_by(state_po) %>% 
  summarize(n = n(),
          m_totalvotes = mean(totalvotes),
          sd_totalvotes = sd(totalvotes),
          mean_pop = mean(pop), 
          sd_pop = sd(pop)))

(table_02 <- df %>% 
  group_by(year_elctn, party) %>% 
  summarize(n = n(),
          m_candidatevotes = mean(candidatevotes),
          sd_candidatevotes = sd(candidatevotes)))

```


```{r descriptives tables}

(table_1 <- table_01 %>% 
  kbl(col.names = c("Sate", "N", "M", "SD",
                    "M", "SD"), # renaming the columns
      digits = 2,
      caption = "<b>Table 1</b><br /> <i>A summary table for total votes and population by state.</i>",
      align =  "c") %>% #align center 
  kable_classic(full_width = FALSE) %>% 
  add_header_above(c(" " = 2, "Total Votes" = 2, "Population" = 2)) %>% #adding headers. The 2's indicate how many columns are under each header
  footnote(footnote_as_chunk = TRUE, 
           general = "N = number of counties with recorded data per the four time points (2000, 2008, 2012, 2016).")) #adding a footnote

(table_2 <- table_02 %>% 
  kbl(col.names = c("Year", "Party", "N", "Mean Candidate Votes", "SD Candidate Votes"),
      digits = 2,
      caption = "<b>Table 2</b><br /> <i>A summary table for votes by candidate and year of election.</i>",
      align =  "c") %>%
  kable_classic(full_width = FALSE) %>% 
  footnote(footnote_as_chunk = TRUE, 
           general = "N = total number of counties in the US reporting data.")) 

```



``` {r preliminary analysis}
df2 <-df  %>% 
  pivot_wider(
    names_from = party,
    values_from = candidatevotes) %>% 
  mutate(demmargin = (democrat - republican) / (democrat + republican))

# social capital variables with per capita measures
df_anal_2016 <- df2 %>%
  filter(year_elctn == 2016) %>% 
  select(bowl:demmargin) %>%
  select(-democrat, -republican, -green, -"NA", -sk) %>% 
  mutate(bowl =bowl/pop * 1000) %>%
  mutate(civic = civic/pop * 1000) %>% 
  mutate(golf = golf/pop * 1000) %>% 
  mutate(relig = relig/pop * 1000) %>% 
  mutate(sport = sport/pop * 1000) %>% 
  mutate(pol = pol/pop * 1000) %>%
  mutate(prof = prof/pop * 1000) %>% 
  mutate(bus = bus/pop * 1000) %>% 
  mutate(labor = labor/pop * 1000) %>%
  mutate(nccs = nccs /pop * 1000)

df_anal_2016 %>% 
  cor()

#preliminary analysis
mod1 <- lm(data = df_anal_2016, demmargin ~ 1 + bowl + civic + golf + relig + sport + pol + prof + bus + labor + pvote + respn  + pop)
summary(mod1)

```


```{r}
#Scatterplot of social capital over time (individual points represent states). Line fitted to show any trends over time
aggregate_states <- df %>%
  filter((year_elctn == 2000 & party == "republican") | 
           (year_elctn == 2008 & party == "democrat") | 
           (year_elctn == 2012 & party == "democrat") | 
           (year_elctn == 2016 & party == "republican")) %>%
  mutate(aggregate_capital = select(., bowl:labor) %>% rowSums()) %>%
  group_by(state_po,year_sc) %>%
  summarize(aggregate_capital=sum(aggregate_capital))
aggregate_states %>%
  ggplot(mapping=aes(x=year_sc,y=aggregate_capital)) +
  geom_point() +
  geom_smooth(se=FALSE,method=lm) +
  labs(title="Social Capital Trends in U.S. States",
       x="Year",
       y="Aggregate Social Capital")

#Scatterplot of different types of social capital over time (individual points represent aggregate for entire country). Lines fitted for each type to show trends over time
filter_df <- df %>%
    filter((year_elctn == 2000 & party == "republican") | 
           (year_elctn == 2008 & party == "democrat") | 
           (year_elctn == 2012 & party == "democrat") | 
           (year_elctn == 2016 & party == "republican")) 
aggregate_country <- rowsum(filter_df[9:17], filter_df$year_sc)
aggregate_country <- aggregate_country %>%
  mutate(Year = c(1997,2005,2009,2014)) %>%
  pivot_longer(cols=bowl:labor,names_to = "type_sc", values_to = "count")
aggregate_country %>%
  ggplot(mapping=aes(x=Year,y=count,color=type_sc)) +
  geom_point() +
  geom_smooth(se=FALSE,method=lm) +
  labs(title="Changes in Social Capital in U.S.",
       x="Year",
       y="Count",
       color="Type of Capital")

#Scatterplot of proportion votes for winning party by social capital for each state
aggregate_election <- df %>%
  filter((year_elctn == 2000 & party == "republican") | 
           (year_elctn == 2008 & party == "democrat") | 
           (year_elctn == 2012 & party == "democrat") | 
           (year_elctn == 2016 & party == "republican")) %>%
  mutate(aggregate_capital = select(., bowl:labor) %>% rowSums()) %>%
  group_by(state_po,year_elctn,party) %>%
  summarize(aggregate_capital=sum(aggregate_capital), 
            prop_votes = sum(candidatevotes) / sum(totalvotes))
aggregate_election %>%
  ggplot(mapping=aes(x=aggregate_capital,y=prop_votes,color=party)) +
  geom_point() +
  geom_smooth(se=FALSE,method=lm,color="black") +
  facet_wrap(~year_elctn) +
  labs(title="Votes for Winning Party by Social Capital in U.S. States",
       x="Aggregate Social Capital",
       y="Proportion of Votes",
       color="Winning Party")

  
```

```{r visualization part 2}
#install.packages("usmap")
library(usmap)

#percentage of democratic voters
df2014 <- df2 %>% 
  filter(year_elctn == 2016)

#US Election Results 2016
plot_usmap(data = df2014, regions = "states", values = "demmargin", labels = TRUE, label_color = "white") +
  scale_fill_gradient2(low = "red", 
                       mid = "white", 
                       midpoint = 0, 
                       high = "blue", 
                       name = "Republican - Democrat Margin (2016)") + 
  #scale_fill_continuous(low = "red", high = "blue", name = "Percentage of Democratic Votes (2016)", label = scales::comma) + 
  theme(panel.background = element_rect(color = "black", fill = "lightblue")) + 
  theme(legend.position = "bottom") 

df2008 <- df2 %>% 
  filter(year_elctn == 2008)

plot_usmap(data = df2008, regions = "counties", values = "demmargin") +
  scale_fill_gradient2(low = "red", 
                       mid = "white", 
                       midpoint = 0, 
                       high = "blue", 
                       name = "Republican - Democrat Margin (2008)") + 
  #scale_fill_continuous(low = "red", high = "blue", name = "Percentage of Democratic Votes (2016)", label = scales::comma) + 
  theme(panel.background = element_rect(color = "black", fill = "lightblue")) + 
  theme(legend.position = "bottom")


df_usmap <- df2 %>% 
  filter(state_po != "AL")

unique(df_usmap$year_elctn)

plot_usmap(data = df_usmap, regions = "counties", values = "demmargin") +
  scale_fill_gradient2(low = "red", 
                       mid = "white", 
                       midpoint = 0, 
                       high = "blue", 
                       name = "Republican - Democrat Margin (2016)") + 
  #scale_fill_continuous(low = "red", high = "blue", name = "Percentage of Democratic Votes (2016)", label = scales::comma) + 
  theme(panel.background = element_rect(color = "black", fill = "lightblue")) + 
  theme(legend.position = "bottom") + 
  facet_wrap(~year_elctn, ncol = 2)



#Pipes don't really work with usmap package...





#Oregon election 016
plot_usmap(data = df2014, regions = "counties", include = c("OR"), values = "demmargin", labels = TRUE) +
    scale_fill_gradient2(low = "red", 
                         mid = "white", 
                         midpoint = 0, 
                         high = "blue", 
                         name = "Republican - Democrat Margin (2016)") +
  theme(panel.background = element_rect(color = "black", fill = "lightblue")) + 
  theme(legend.position = "bottom")

#US Social Capital 2014
plot_usmap(data = df2014, regions = "counties", values = "assn") +
  scale_fill_continuous(low = "white", high = "blue", name = "Social Capital Index (2014)", label = scales::comma
  ) + 
  theme(panel.background = element_rect(color = "black", fill = "lightblue")) + 
  theme(legend.position = "bottom")


#Oregon social capital 2014
plot_usmap(data = df2014, regions = "counties", include = c("OR"), values = "assn") +
  scale_fill_continuous(low = "white", high = "blue", name = "Social Capital Index (2014)", label = scales::comma
  ) + 
  theme(panel.background = element_rect(color = "black", fill = "lightblue")) + 
  theme(legend.position = "bottom")


```

# Results

# Discussion


\newpage

# References

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id="refs" custom-style="Bibliography"></div>
\endgroup
