---
title             : "Relationship between social capital and election results"
shorttitle        : "Title"

author: 
  - name          : "Anisha Babu"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    email         : "ababu@uoregon.edu"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - Conceptualization
      - Data Analysis
      - Writing - Original Draft Preparation
      - Writing - Review & Editing
      
  - name          : "Hyeonjin Cha"
    affiliation   : "1"
    email         : "hcha@uoregon.edu"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - Conceptualization
      - Data Analysis
      - Writing - Original Draft Preparation
      - Writing - Review & Editing
      
  - name          : "Diana DeWald"
    affiliation   : "1"
    email         : "ddewald@uoregon.edu"    
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - Conceptualization
      - Data Analysis
      - Writing - Original Draft Preparation
      - Writing - Review & Editing
      
  - name          : "Murat Kezer"
    affiliation   : "1"
    email         : "mkezer@uoregon.edu"    
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - Conceptualization
      - Data Analysis
      - Writing - Original Draft Preparation
      - Writing - Review & Editing               

affiliation:
  - id            : "1"
    institution   : "University of Oregon"


authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

  Enter author note here.

abstract: |
  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.
  
  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.
  
  One sentence clearly stating the **general problem** being addressed by  this particular study.
  
  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).
  
  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.
  
  One or two sentences to put the results into a more **general context**.
  
  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.
  
  <!-- https://tinyurl.com/ybremelq -->
  
keywords          : "keywords"
wordcount         : "X"

bibliography      : ["r-references.bib"]

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
library("papaja")
r_refs("r-references.bib")
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed,
                      eval = TRUE,
                      echo = FALSE,
                      cache = FALSE)
```

```{r packages}
library(rio); library(here); library(tidyverse); library(knitr); library(janitor); library(magrittr); library(stringr)
```

# Data Preparation
```{r load data}
election <- import(here("data", "countypres_2000-2016.csv")) %>% 
  clean_names() #load election results

s_capital_97 <- import(here("data", "social-capital-variables-for-1997.xlsx")) %>% 
  clean_names() #load social capital data for 1997

s_capital_05 <- import(here("data", "social-capital-variables-for-2005.xlsx")) %>% 
  clean_names() #load social capital data for 2005

s_capital_09 <- import(here("data", "social-capital-variables-for-2009.xlsx")) %>% 
  clean_names() #load social capital data for 2009

s_capital_14 <- import(here("data", "social-capital-variables-spreadsheet-for-2014.xlsx")) %>% 
  clean_names()#load social capital data for 2014
```

```{r clean datasets}
#select, filter, and clean data
##election
election %<>%
  select(-office, -version, -candidate, -state) %>% #select relevant variables
  filter(year %in% c(2000, 2008, 2012, 2016)) #select relevant years

#social capital 1997
s_capital_97[290, 2] <- "District of Columbia, DC" #add DC's code

s_capital_97 %<>% #create county and state variables
  separate(areaname, c("county","state_po"), sep = ', ')

s_capital_97 %<>% #select relevant variables
  select(-fitnes97, -memspt97, -memnec97)

colnames(s_capital_97)[4:18] <- str_sub(colnames(s_capital_97)[4:18], end=-3) #remove the year from the variable names

s_capital_97$year <- 1997 #create year variable before merging

#social capital 2005
s_capital_05[290, 2] <- "District of Columbia, DC" #add DC's code

s_capital_05 %<>% #create county and state variables
  separate(areaname, c("county","state_po"), sep = ', ')

s_capital_05 %<>% #select relevant variables
  select(-fitns05)

colnames(s_capital_05)[4:18] <- str_sub(colnames(s_capital_05)[4:18], end=-3) #remove the year from the variable names

s_capital_05 <- s_capital_05[,c(1:3, 10, 5, 11, 4, 12, 7:8, 6, 9, 15, 14, 13, 16, 17:18)] #reorder the variables

s_capital_05$year <- 2005 #create year variable before merging

#social capital 2009
s_capital_09[290, 2] <- "District of Columbia, DC" #add DC's code

s_capital_09 %<>% #create county and state variables
  separate(areaname, c("county","state_po"), sep = ', ')

s_capital_09 %<>% #select relevant variables
  select(-fitns09)

colnames(s_capital_09)[4:18] <- str_sub(colnames(s_capital_09)[4:18], end=-3) #remove the year from the variable names

s_capital_09 <- s_capital_09[,c(1:3, 10, 5, 11, 4, 12, 7:8, 6, 9, 14:15, 13, 16:18)] #reorder the variables

s_capital_09$year <- 2009 #create year variable before merging

#social capital 2014
s_capital_14[77, 2] <- "Hoonah-Angoon Census Area, AK" #remove the repeated state code

s_capital_14 %<>% #create county and state variables
  separate(county_name, c("county","state_po"), sep = ', ') 

s_capital_14$county <- word(s_capital_14$county, 1) #remove the word 'county' and keep only the name

s_capital_14 %<>% #select relevant variables
  select(-recreational2014)

colnames(s_capital_14)[4:18] <- str_sub(colnames(s_capital_14)[4:18], end=-5) #remove the year from the variable names

s_capital_14 <- s_capital_14[, c(1:3, 10, 5, 11, 4, 12, 7:8, 6, 9, 16, 15, 13, 17, 14, 18)] #reorder the variables

s_capital_14$year <- 2014 #create year variable before merging

colnames(s_capital_14) <- colnames(s_capital_09) #rename variables to make it compatible with previous years' datasets
```



# Introduction
Social science literature has extensively examined the relationship between social capital and politics (e.g. Morales & Guigni, 2016; Jottier & Heyndels, 2012; La Due Lake & Huckfeldt, 1998). However, relatively little is known on the impact of social capital election results.

# Methods
We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study. <!-- 21-word solution (Simmons, Nelson & Simonsohn, 2012; retrieved from http://ssrn.com/abstract=2160588) -->

## Participants

## Material

## Procedure

## Data analysis
We used `r cite_r("r-references.bib")` for all our analyses.


# Results

# Discussion


\newpage

# References

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id="refs" custom-style="Bibliography"></div>
\endgroup
